---
title: 'Lab 2: What Makes a Movie Successful? Fall 2021'
author: 'w203: Statistics for Data Science'
date: "November 1, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggplot2) 
library(jsonlite)

library(sandwich)
library(stargazer)
```

## Introduction

<!-- Possibly add more about why the models are constructed the way they are --> 

What makes a movie “one of the greats”? While that question may elude both data analysts and artists for years to come, we are looking to find some effective indicators of box office success for films, in the hopes of maintaining a thriving movie industry for years to come. This project will focus on several commonly cited reasons for movie fiscal success and analyze how much of an impact each factor has. 

The client for this analysis could be any major movie studio or its parent company, e.g. NBCUniversal or Walt Disney Studios. This is a significant issue, because box office sales decreased by 60% in 2020 during COVID-19 pandemic, leaving the movie theaters empty for months. This analysis will provide clues for what a studio can do to speed up the recovery and account for lost revenue.

The mmain dataset we’ll use for this analysis comes from The Movie Database (TMDB - https://www.themoviedb.org/), as well as several assembled lists from online polls (e.g. “Greatest Actors/Actresses of All Time” - Ranker.com). 

We’re planning on using profitability as the response variable for analysis in this study - using R to clean the empty and irrelevant cells, a “profitability ratio” can be determined for each movie based on its box office sales divided by budget (and controlling for foreign films and currencies).

In order to build our models, we decided to focus on some factors that are usually linked to a movie's financial success, such as budget, time of release, its cast's popularity and director's name brand recognition. We will also take into account some control variables such as runtime and genre.

In assessing the value of name-brand recognition in Hollywood, we’ll use R and its extensions to cross-check the presence of the “greatest” actors and actresses of all time, assigning a count to each movie to later plot against profitability. We’ll be using an Ordinary Least Squares regression model to predict movie success as a factor of famous actor/director presence, online ratings, and other variables that we find significant.

Once we obtain an accurate model, we can obtain numbers such as the monetary value of a single point increase in rating, to explain the trends and interpret results. We can consider several situations accounting for premiere dates, return on investment (ROI), runtime along variables mentioned earlier. While there are certain limitations to this data set and analysis e.g. limited number of data points, some data entry errors,  inaccurate labeling and biased focus on foreign films, and subjective descriptions of the movies, this should serve as a tool and starting point for aspiring and current filmmakers.

The products of this analysis will include the code and final datasets, as well as a written summary of the significant results. 


```{r setup, include=FALSE}
d <- read.csv('./src/data/Clean_Data_set.csv')

actors_df <- read.csv('./src/data/Greatest Actors List.csv')

directors_df <- read.csv('./src/data/Greatest Directors List.csv')
```


```{r setup, include=FALSE}
# exclude unnecessary variables
excluded_variables <- names(d) %in% c("homepage", "overview", "imdb_id", "popularity", "poster_path", "status", "tagline", "title", "Keywords", "production_companies", "production_countries", "spoken_languages", "cast", "crew")
short_data <- d[!excluded_variables]

#remove empty actors entry
actors_df<- actors_df[(actors_df$Name != ''),]

#rename director columns
names(directors_df)[names(directors_df) == 'Name'] <- 'director_name'
names(directors_df)[names(directors_df) == 'Score'] <- 'director_score'
names(short_data)[names(short_data) == 'directors'] <- 'director_name'


#drop unnecessary director columns
excluded_variables <- names(directors_df) %in% c("Rank", "Count", "M.or.F")
directors_df <- directors_df[!excluded_variables]
```

```{r clean data, include=FALSE}
# Remove entries without budget or genre or cast
clean_data<-short_data[(short_data$budget > 0),]
clean_data<-clean_data[(clean_data$actors != "[]"),]
# Convert belongs_to_collection to a True or False variable
clean_data$belongs_to_collection[clean_data$belongs_to_collection != ""] <- TRUE
clean_data$belongs_to_collection[clean_data$belongs_to_collection == ""] <- FALSE

```


```{r add actor score, include=FALSE}
top_actors <- actors_df$Name
count_top_actors <- function(list) {
  counter <- 0
  for (actor in top_actors) {
    if (regexpr(actor, list[11]) >= 0) {
      counter <- counter + 1
    }
  }
  return(counter)
}

clean_data$actor_score <- apply(clean_data, 1, FUN = count_top_actors)
```

```{r add director score, include=FALSE}
merged_data <- merge(clean_data, directors_df, by = "director_name", all.x = TRUE)
merged_data$director_score[is.na(merged_data$director_score)] <- 0
```

```{r}
final_table <- merged_data %>% 
  mutate(
    comedy = case_when(
      #encode comedy, and all subsequent genres as binary
      regexpr('Comedy', genres) >= 0 ~ 1, 
      regexpr('Comedy', genres) == -1 ~ 0),
    
    drama = case_when(
      regexpr('Drama', genres) >= 0 ~ 1, 
      regexpr('Drama', genres) == -1 ~ 0),

    family = case_when(
      regexpr('Family', genres) >= 0 ~ 1, 
      regexpr('Family', genres) == -1 ~ 0),

    romance = case_when(
      regexpr('Romance', genres) >= 0 ~ 1, 
      regexpr('Romance', genres) == -1 ~ 0), 
    
    thriller = case_when(
      regexpr('Thriller', genres) >= 0 ~ 1, 
      regexpr('Thriller', genres) == -1 ~ 0), 
    
    animation = case_when(
      regexpr('Animation', genres) >= 0 ~ 1, 
      regexpr('Animation', genres) == -1 ~ 0), 
    
    adventure = case_when(
      regexpr('Adventure', genres) >= 0 ~ 1, 
      regexpr('Adventure', genres) == -1 ~ 0), 
    
    horror = case_when(
      regexpr('Horror', genres) >= 0 ~ 1, 
      regexpr('Horror', genres) == -1 ~ 0), 
    
    music = case_when(
      regexpr('Music', genres) >= 0 ~ 1, 
      regexpr('Music', genres) == -1 ~ 0), 
    
    crime = case_when(
      regexpr('Crime', genres) >= 0 ~ 1, 
      regexpr('Crime', genres) == -1 ~ 0), 
    
    sci_fi = case_when(
      regexpr('Science Fiction', genres) >= 0 ~ 1, 
      regexpr('Science Fiction', genres) == -1 ~ 0), 
    
    action = case_when(
      regexpr('Action', genres) >= 0 ~ 1, 
      regexpr('Action', genres) == -1 ~ 0), 
    
    war = case_when(
      regexpr('War', genres) >= 0 ~ 1, 
      regexpr('War', genres) == -1 ~ 0), 
    
    western = case_when(
      regexpr('Western', genres) >= 0 ~ 1, 
      regexpr('Western', genres) == -1 ~ 0), 
    
    fantasy = case_when(
      regexpr('Fantasy', genres) >= 0 ~ 1, 
      regexpr('Fantasy', genres) == -1 ~ 0), 
    
    foreign = case_when(
      regexpr('Foreign', genres) >= 0 ~ 1, 
      regexpr('Foreign', genres) == -1 ~ 0), 
        
    mystery = case_when(
      regexpr('Mystery', genres) >= 0 ~ 1, 
      regexpr('Mystery', genres) == -1 ~ 0), 
    
    history = case_when(
      regexpr('History', genres) >= 0 ~ 1, 
      regexpr('History', genres) == -1 ~ 0), 
    
    documentary = case_when(
      regexpr('Documentary', genres) >= 0 ~ 1, 
      regexpr('Documentary', genres) == -1 ~ 0),
    
    january = case_when(
      
      substr(release_date, start = 1, stop = 2) == '1/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '1/' ~ 0),
    
    
     february = case_when(
      
      substr(release_date, start = 1, stop = 2) == '2/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '2/' ~ 0),
    
    march = case_when(
      
      substr(release_date, start = 1, stop = 2) == '3/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '3/' ~ 0),
    
    april = case_when(
      
      substr(release_date, start = 1, stop = 2) == '4/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '4/' ~ 0),
    
    may = case_when(
      
      substr(release_date, start = 1, stop = 2) == '5/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '5/' ~ 0),
       
     june = case_when(
      
      substr(release_date, start = 1, stop = 2) == '6/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '6/' ~ 0),
    
    july = case_when(
      
      substr(release_date, start = 1, stop = 2) == '7/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '7/' ~ 0),
    
    
    august = case_when(
      
      substr(release_date, start = 1, stop = 2) == '8/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '8/' ~ 0),
    
    september = case_when(
      
      substr(release_date, start = 1, stop = 2) == '9/' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '9/' ~ 0),
    
    october = case_when(
      
      substr(release_date, start = 1, stop = 2) == '10' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '10' ~ 0),
    
    november = case_when(
      
      substr(release_date, start = 1, stop = 2) == '11' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '11' ~ 0),
    
    december = case_when(
      
      substr(release_date, start = 1, stop = 2) == '12' ~ 1, 
      substr(release_date, start = 1, stop = 2) != '12' ~ 0)
    
    
  )
      

```




```{r}

ggplot(final_table, aes(x=budget)) + geom_histogram(bins = 100)

ggplot(final_table, aes(x=revenue)) + geom_histogram(bins = 100)


# The logarithm of both budget and revenue appear more homogenously distributed
ggplot(final_table, aes(x=log(budget))) + geom_histogram(bins = 100)

ggplot(final_table, aes(x=log(revenue))) + geom_histogram(bins = 100)

ggplot(final_table, aes(x=runtime)) + geom_histogram(bins = 100)
#runtime appears to be relatively normally distributed overall, except for the narrow group on the lower end of time


#ggplot(final_table, aes(x=release_date)) + geom_bar()


ggplot(final_table, aes(x=substr(release_date, start = 1, stop = 2))) + geom_bar()
#Release date overall looks fairly well distributed over time - the number indicates month, with 1/ - 9/ representing  January to September, and 10 - 12 representing October - December


```

```{r}

plot(log(final_table$budget), log(final_table$revenue))
title(main = 'Logarithmic Movie Revenue over Budget', xlab = 'Log of Budget', ylab = 'Log of Revenue')
# The shape of the relationship looks to be positively correlated, and seems to suggest that a higher budget results in a higher revenue

plot(final_table$runtime, log(final_table$revenue))
title(main = 'Logarithmic Movie Revenue over Movie Runtime', xlab = 'Runtime', ylab = 'Log of Revenue')

```



## Research Question, Underlying Model and Study Design

### Research Question

This study intends to find significant relationships between factors known before a movie release and its box office success. We will look at both its profitability ratio (defined as box office divided by its budget) and its box office as ways of measuring a movie's success.

The results of this study could be used by film makers to better inform their investment decisions.

### Underlying Model

As mentioned in the introduction, we plan on using commonly cited reasons for movie's financial success to create our models, starting with a simplified model with just the key variables, and then add more variables we hypothesize could have an effect on a movie's financial success to the following models

#### Simplified model

The most simple and straight forward way of predicting a movie's box office would be its budget. The relationship is pretty clear. The more money that is put into a project the better it is expected to do. One would also expect the increase budget would go into improving the quality of the movie that in turn should increase its box office performance.

```{r}

ggplot(final_table, aes(x=budget/1000000, y=revenue/1000000)) + geom_point() + labs(
  title = 'Movie Revenue over Budget',
  x = "Budget in MM USD",
  y = "Revenue in MM USD"
)

ggplot(final_table, aes(x=log(budget), y=log(revenue))) + geom_point() + labs(
  title = 'Logarithmic Movie Revenue over Budget',
  x = "Log of. Budget in USD",
  y = "Log of Revenue in USD"
)

```

As seen in the previous charts, there is a clear positive relationship between budget and revenue, both for with and without a logarithmic transformation. This contributes to our hypothesis than a higher budget leads to a better box office.

#### Seasonality

Just like for a lot of other industries, we expect for there to be a seasonal effect on a movie's box office success. We theorize that the effect of seasonality in the movie industry comes from both the demand and the supply.

Certain times of the year tend to attract more moviegoers, such as the summer, the holiday season and certain weekends that coincide with important holidays (Christmas, 4th of July weekend, Memorial Day weekend, among others). This increase demand would lead to a better box office performance.

On the other hand, supply is also affected by seasonality. Big budget movies are more likely to be released during a time of the year that could help them maximize their potential revenue. 

This self-selection on the supply side is likely to have an effect on the demand as well, as a greater number of quality movies during certain times would attract more movie goers.

Our model will control for this effect by adding the month of release to the model.

#### The Cast's Name Recognition

It shouldn't be a surprise to anyone that famous actors attract crowds, which in turn increase the movie's box office performance. Audiences want to see their favorite actors in their movies, for a variety of reasons.

A recent example that comes to mind is the 2021 movie adaptation of Dune. The actress Zendaya was a big part of the movie's marketing, despite the fact that she was only briefly appears in the movie. A lot of movie goers were upset after seeing the movie since they expected her to have a bigger role.

For our model we will give each movie an "Actor Score" which is just a count of how many well known actors are part of their cast.


### Research design

<!-- Type of research design --> 
<!-- Type of models and goals --> 

### Data

As mentioned before our main dataset comes from data extracted from The Movie Database (TMDB - https://www.themoviedb.org/). This dataset was compiled by Kaggle for a Machine Learning competition a few years ago.

The original dataset consists of 22 variables for over 7000 films, from the 1920s to 2017. This dataset consists of objective data for each of these movies such as budget, revenue, cast, crew, release date, production company, original language among others.

We also added two more datasets consisting of two lists with a ranking of the "Top 100 movie actors of all time", as well as the top "Top 50 directors". Unlike the previous dataset, these ranks is subjective, but we plan on using these lists to help us create a variable that could stand in for an actor's or director's popularity.

We performed some data cleaning operations on the main dataset in order to have it be ready for our analysis. These included:

- Eliminating any entries that had some vital information missing (such as revenue, budget, cast, crew and genres)
- Parsing certain fields from a stringified JSON.
- Creating the "Actors" column from the "Cast" column
- Extracting the Director from the "Crew" column

After completing this process we still had over 5000 entries.

## Statistical Model

<!-- EDA  --> 
<!-- At least three model specifications: one with key variables, two defensible models  --> 

<!-- What are we measuring --> 
<!-- Covariates, problems with covariates --> 
<!-- Transformations --> 

## Results

```{r regression table, include=FALSE}
# stargazer(
#   mod_pooled_no_covariates,
#   mod_pooled_with_covariates,
#   type = 'latex', header = FALSE,
#   star.cuttoffs = c(0.05, 0.01, 0.001)
# )
```
<!-- Statistical significance and practical significance --> 
<!-- Other stat tests --> 
## Limitations
### Statistical limitations
<!-- Evaluate problematic large model assumptions --> 
### Structural limitations
<!-- Omitted variables --> 
## Conclusion and Discussion 

<!-- Key insights --> 
